Joint Likelihood Deconvolution of Astronomical Images in the Presence of Poisson Noise

### 2023-05-08 src/tex/ms.tex
[showyourwork] overleaf sync

% Abstract with filler text
    \begin{abstract}
        We present a new method for Joint Likelihood Deconvolution (Jolideco) of a set of astronomical observations of the same sky region in the presence of Poisson noise. The method reconstructs a single flux image from a set of observations by optimizing the \aposteriori joint Poisson likelihood of all observations under a patch based image prior. The patch prior is parameterised by a Gaussian Mixture model (GMM) and learned from astronomical images with high signal to noise ratio, such as the James Webb Telescope. During the reconstruction process the patch prior adapts  to the structures in the data by finding the most likely GMM component for each patch in the image. By applying the method to simulated data we show that both the combination of multiple observations as well as the patch based prior lead to a much improved reconstruction quality in many different source scenarios as well as signal to noise regimes. We show also that the method yields superior reconstruction quality to alternative standard methods such as the Richardson-Lucy method. We also show that the method can be highly adapted to various real analysis scenarios by learning problem specific GMM prior  such as images of extragalactic or Galactic sources.
    \end{abstract}

@@

    \newpage
    \subsection{Related Work}
    In literature there have been many efforts to solve the unblind deconvolution problem for images affected by Poisson noise. Most methods rely fundamentally on modeling the image generating process as a linear process. Given a known point spread function a prediction of the discrete measured image can be obtained via \textit{forward folding} and be compared against an observation assuming Poisson statistics per pixel. However the methods differ in their additional prior or regularisation assumptions as well as their approach to optimization or sampling of the underlying likelihood function.

@@

    However in practice the RL method was found to have multiple, partly related limitations: first there is no objective criterion when to stop the optimization process. Usually the iteration is stopped when the reconstructed image seems plausible to the eye. Second, when the number of iterations is too high, the method shows a tendency to decompose regions of constant brightness into a set of bright emission peaks (\textit{speckles}). This problem was e.g. described and studied by \cite{Reeves1995} and \cite{Fish1995}. Lastly the RL method does not provided an associated estimate of the uncertainties of the reconstruction, which is desirable especially in the Poisson domain to measure the significance of specific features found in the reconstructed image.

@@

    To address the issue with RL decomposing extended emission structures into groups of point source, multiple additions to the RL method were proposed. Most importantly this lead to the introduction of regularised RL methods.
    such as use of total variation \citep{Dey2006} or sparseness \todo{add reference}.

@@

    An alternative approach to solve the problem in the framework of information field theory. Recently, image deconvolution using physical assumptions and decomposition into point and diffuse flux in 3D/4D has gained attention. \cite{Pumpe2018} and \cite{Selig2015} proposed priors based on physical assumptions, which have been shown to work well for large-scale emission. However, the smoothness assumption in these methods can lead to artifacts in the galactic plane.

    The challenge is due to fundamentally contradicting goals: deconvolution favoring point sources will decompose extend structures, while introducing local correlations will lead to

    \cite{Zoran2011} build on very similar assumption for natural images.
    Instead of assuming a fixed local correlation structures, such as smoothness
    they proposed to learn the local correlation structure on millions of patches
    extracted from natural images.

    For this they proposed to split training images from representative
    data sets into small patches of size 8x8~pixels and fit a 64 dimensional
    Gaussian Mixture Model (GMM) to the distribution of extracted patches. Each pixel of the patch 
    is thereby treated as an independent dimension in the model. In total
    they chose to use $k=200$ Gaussian components.

    To use the GMM in reconstruction they introduce the expected patch likelihood (EPLL).

    They showed that the patch based GMM prior led to much improved image reconstructions for multiple inverse problems such as denoising, deblurring and inpainting.

    \cite{Bouman2016} later adapted the patch prior reconstruction to be used

@@

    Some of the challenges of deblurring with CNN are due to the convolutional nature of the networks. In a single layer, linear approximation, the network would need to learn the inverse filter of the PSF, corresponding to the well known \textit{Wiener filter} solution of the problem. This requires large convolution kernels or deep network architectures. \cite{Li2014} have introduced a baseline architecture for a deep CNN for image deblurring by introducing large kernels under the separability assumption.

    One last important challenge is obtaining uncertainty estimates for the reconstructed image from CNNs. In the limit of Poisson noise it is desirable to trace the full a-postiori likelihood, which is not possible for most network architectures, with the exception of for example normalising flows.

@@

    For each individual observation $j$ the predicted counts can be modelled by \textit{forward folding} the unknown flux image $\mathbf{x}$ through the individual instrument response:

    \begin{equation}
        \label{eq:model}
        \boldsymbol{\lambda}_j(\mathbf{x}) = \mathrm{PSF}_j \circledast \left(\mathbf{E}_j \cdot (\mathbf{x} + \mathbf{B}_j) \right)
    \end{equation}

@@

    Given a single observation $j$ of an unknown flux image
    $\mathbf{x}$ and assuming the noise in each pixel $i$ in the recorded counts image
    $\mathbf{D_j}$ follows a Poisson distribution, the likelihood $\mathcal{P}$
    of obtaining the measured image from a model image of the expected
    counts $\boldsymbol{\lambda_j}$ with $N$ pixels is given by:

    \begin{equation}
        \label{eq:poisson}
        P\left( \mathbf{D_j} | \boldsymbol{\lambda_j} \right) = \prod_{i=1}^N \frac{{e^{ - D_{j,i} } \lambda_{i,j} ^ {D_{j,i}}}}{{D_{j,i}!}}
    \end{equation}

    By taking the logarithm and dropping the constant terms one can transform the
    product into a sum over pixels, which is also often called the \textit{Cash}
    \citep{Cash1979} fit statistics:

    \begin{equation}
        \label{eq:cash}
        \mathcal{C}\left( \mathbf{D_j} | \boldsymbol{\lambda_j} \right) = \frac{1}{N}\sum_{i=1}^N \lambda_{j,i} - D_{j, i} \log{\lambda_{j,i}}
    \end{equation}

    To account for the possibility of different datasets having a different number of pixels we divide by $N$ (\todo{Does this need more justification...?}).
    By summing over all observations we finally get the joint log-likelihood of measuring a set of counts images $\mathbf{D}$ given an unknown flux image $\mathbf{x}$:

    \begin{equation}
        \label{eq:joint}
        \mathcal{L}\left( \mathbf{D} | \mathbf{x} \right) = \sum_{j=1}^J \mathcal{C}\left( \mathbf{D_j} | \mathbf{x} \right)
    \end{equation}

@@

    Using Bayes rule we can estimate the \aposteriori likelihood under a given prior:

    \begin{equation}
        \label{eq:bayes}
        P(\mathbf{x}|\textbf{D}) = P(\mathbf{x} ) \frac{P(\textbf{D} |\mathbf{x})}{P(\textbf{D})}
    \end{equation}

    And use a Maximum A Posteriori (MAP) approach to find an estimate for $\mathbf{x}$.
    Taking the logarithm of Eq.~\ref{eq:bayes}, replacing the definition of the likelihood
    $P(\mathbf{D}|\mathbf{x})$ with Eq.~\ref{eq:joint} and dropping the normalisation term $P(\mathbf{D})$
    which is independent of $\mathbf{x}$, we get the following expression for the log-posterior
    likelihood $\mathcal{L}$ :

    \begin{equation}
        \label{eq:total}
        \mathcal{L}\left(\mathbf{x} | \mathbf{D} \right) = \sum_{j=1}^J \mathcal{C}\left( \mathbf{D}_j | \mathbf{x} \right) - \beta \cdot \mathcal{P}(\mathbf{x})
    \end{equation}

@@

    \subsection{Patch Based Image Prior}
    In general it is difficult to capture global image statistics in a prior likelihood taking into account all pixel to pixel correlations. The size of the corresponding correlation matrix would grow with the number of pixels squared and thus quickly becomes computationally untraceable. However astronomical images typically only show local correlations between pixels. This is because the spatial scale of astrophysical processes in an image is always limited by the finite interaction time and the fact that the interaction cannot propagate faster than the speed of light (structure in the CMB might be an exception here). This means on smaller scales astronomical images often contain basic structures such as edges, corners, filaments or periodic patterns formed by local astrophysical processes. For this reason we adopt the idea of a patch based image prior approach as introduced by \cite{Zoran2011}. We use a pre-trained Gaussian Mixture Model (GMM) to evaluate and approximate a patch based prior for the reconstructed image $\mathbf{x}$.

    One convenient property of the GMM is the possibility to evaluate its likelihood in closed form. For a given single patch $\mathbf{x}_n$ it is given by:

    \begin{equation}
        \label{eq:gmm}
        P_{GMM}(\mathbf{x}_n) = \sum_{k=1}^K \pi_k \mathcal{N}(\mathbf{x}_n| \boldsymbol{\mu}_k, \boldsymbol{\Sigma}_k^2)
    \end{equation}

@@

    From Equation~\ref{eq:gmm} we can also derive the log-likelihood for a specific choice of the component $k$:

    \begin{equation}
    \begin{split}
    \mathcal{P}_{GMM}(k| \mathbf{x}_n) = -\frac{1}{2} \left[ \right. d \log(2\pi)
    + \log(|\boldsymbol \Sigma_{k}|)\\
    + (\mathbf{x}_n - \boldsymbol{\mu}_{k})^T \Sigma^{-1}_{k}(\mathbf{x}_n - \boldsymbol{\mu}_{k}) \left. \right]
    \end{split}
    \end{equation}

    If we split the image into non-overlapping patches, we can find for each noisy patch $\mathbf{x}_n$ the GMM component $\hat{k}_n$, which maximises the log-likelihood of the GMM:

    \begin{equation}
        \hat{k}_n = \underset{k \in \{1, ..., N\} }{\argmax}{\mathcal{P}_{GMM}(k| \mathbf{x}_n)}
    \end{equation}
    \vspace{0.2em}

    As the number of components in the GMM is limited we can do this simply by evaluating the likelihood for each of the components and choose the one with
    the maximum value of the log-likelihood. To finally evaluate the log-likelihood
    of the total image we can sum up the log-likelihood of the individual patches:

    \begin{equation}
        \mathcal{P}(\mathbf{x}, \mathbf{k}) = \sum_{n = 1}^N \mathcal{P}_{GMM}(\hat{k}_n | \mathbf{P}_n \mathbf{x})
    \end{equation}

@@

@@

    To account for both systematic effects we extend the model for the predicted
    counts introduced in Eq.~\ref{eq:model}, by a background normalisation factor
    $\alpha_j$ and a linear shift in position $\phi_j(\mathbf{x}| \delta_{x,j}, \delta_{y,j})$, which is specific for each observation. This yields the following modified model definition:

    \begin{equation}
        \label{eq:model-npred-calibration}
        \mathbf{\lambda}_j = \mathrm{PSF}_j \circledast \left(\mathbf{E}_j \cdot (\phi_j(\mathbf{x}| \delta_{x,j}, \delta_{y,j}) + \alpha_j \cdot \mathbf{B}_j) \right)
    \end{equation}

@@

    \todo{One could try to estimate a sparse covariance matrix...or apprxoimate it using the GMM covariance...}

@@

    We implemented the \jolideco method as an independent Python package, based on \texttt{PyTorch} as optimization back-end. We tried to define a modular, object oriented code structure, to allow parts of the algorithm to be flexible and interchangeable, such as choice of image normalisation scales, the choice of GMM models, optimization methods and serialisation formats for the reconstruction results as well as corresponding diagnostic information such as the trace of the posterior and prior likelihood values. The package is available at \url{https://github.com/jolideco/jolideco}

@@

    \subsubsection{JWST Cassiopeia A}
    The James Webb Space Telescope is the it delivers data a very high resolution and excellent signal to noise ratio. We extracted 2 million patches out of the image...
    Would be 
    Of \textit{Cassiopeia A}
    As a last GM prior we use the promotional image of JWST realeased by...
    The image is particularly suited becasue of the excellent angular resolution, extrating many patches for the same object
    as well as the high signal to noise ratio.

    Taken from \url{https://webbtelescope.org/contents/media/images/2023/121/01GWQBBY77MHGFV3M3N63KDCEJ}

@@

    \begin{figure*}
        \script{example-chandra.py}
        \begin{centering}
            \includegraphics[width=\linewidth]{figures/example-chandra.pdf}
            \caption{
               The \jolideco method applied to \chandra data of the SNR \textit{1E~0102.2–7219}. The left large image shows the summed counts from 25 observations. The right large image  shows the result from the \jolideco deconvolution process. The flux is given as counts relative to the reference observation with ID=1308. The panels \textit{Zoom A} and \textit{Zoom B} show a zoomed-in version of two example regions in the SNR, marked with the white rectangles in the large images. The colormap and scale used in the zoom insets is the same as used for the larger images.
            }
            \label{fig:example-chandra}
        \end{centering}
    \end{figure*}

@@

    The \texttt{fermitools} output an energy dependent counts, exposure and PSF cube. To further reduce the data to images we use the \textit{Gammapy} package. To compute an image of the expected background emission we rely on the latest Galactic diffuse emission model \texttt{gll\_iem\_v07.fits} and also include nearby sources from the 3FHL catalog \cite{Ajello2017}. We first fit the model components to the data using \textit{Gammapy}, excluding a circular region of radius \qty[mode = text]{1.02}{$\deg$} centered on \textit{Vela Junior}. Finally we reduce this model to an image by summing over the energy dimension. To compute the exposure and PSF image we assume a power law with index $\Gamma=1.77$ \citep{Ajello2017} as a spectral model for \textit{Vela Junior} and integrate over the energy dimension of the cube while weighting with the spectral model.

@@

    \section{Paper Reproducibility}
    \label{sec:reproducibility}
    In order to facilitate reproducibility of our results we provide the source code for the paper and figures at \url{https://github.com/jolideco/jolideco-paper}. The paper can be reproduced using using the tool \texttt{Showyourwork} \citep{Luger2021} and following the instructions in the \texttt{README.md} file in the same repository. All data is available for download from Zenodo \todo{how do we do this exactly...?} or can be reproduced using the custom  \texttt{Snakemake} workflows and repositories provided under the \jolideco GitHub organisation \footnote{\url{https://github.com/jolideco}}. For the Fermi-LAT data reduction we used a custom \texttt{Snakemake} workflow  available at \footnote{\url{https://github.com/adonath/snakemake-workflow-fermi-lat}}. For the Chandra data reduction we used a custom \texttt{Snakemake} workflow available at \footnote{\url{https://github.com/adonath/snakemake-workflow-chandra}}.

@@

    To increase the size of the correlation window of the patch prior, \cite{Papyan2015} proposed to evaluate and combine the prior on multiple scales of the image. This is a promising approach to  further improve the reconstruction quality for extended sources with low signal to noise ratio.

    To further improve the expressiveness of the prior \cite{Altekrueger2022} used normalising flows to model the distribution of patches in the context of CT image reconstruction.

    Another promising approach to handle prior information for images has been introduced by \citep{Ulyanov2017} with the so called \enquote{Deep Image Prior}. They showed that a prior structure for an image can be built into a deep neural network. The structure of the network encodes the prior information on correlation between neighbouring pixels and also imposes self similarity on the structures in the image.

>>>>

% Abstract with filler text
    \begin{abstract}
        We present a new method for Joint Likelihood Deconvolution (Jolideco) of a set of astronomical observations of the same sky region in the presence of Poisson noise. The method reconstructs a single flux image from a set of observations by optimizing the \aposteriori joint Poisson likelihood of all observations under a patch based image prior. The patch prior is parameterised by a Gaussian Mixture model (GMM) which we trained on astronomical images with high signal to noise ratio, including data from the  James Webb Telescope as well as the GLEAM radio survey. During the reconstruction process the patch prior adapts to the patch structures in the data by finding the most likely GMM component for each patch in the image. By applying the method to simulated data we show that both the combination of multiple observations as well as the patch based prior lead to a much improved reconstruction quality in many different source scenarios as well as signal to noise regimes. We show also that the method yields superior reconstruction quality to alternative standard methods such as the Richardson-Lucy method. We also the results of the method applied to example data from the \chandra observatory as well as the \fermi. In both cases \jolideco improves the angular resolution as well as signal to noise ratio 
        by a factor of 42 \todo{it would be good to summarize the results with a few key numbers...}.
    \end{abstract}

@@

    \subsection{Related Work}
    In literature there have been many efforts to solve the unblind deconvolution problem for images affected by Poisson noise. Most methods rely fundamentally on modeling the image generating process as a linear process. Given a known point spread function a prediction of the discrete measured image can be obtained via \textit{forward folding} and be compared against an observation assuming Poisson statistics per pixel. However the methods differ in their additional prior or regularisation assumptions as well as their approach to optimization or sampling of the underlying likelihood function.

@@

    However in practice the RL method was found to have multiple, partly related limitations: first there is no objective criterion for when to stop the optimization process. Usually the iteration is stopped when the reconstructed image seems plausible to the eye. Second, when the number of iterations is too high, the method shows a tendency to decompose regions of constant brightness into a set of bright emission peaks (\enquote{speckles}). This problem was e.g. described and studied by \cite{Reeves1995} and \cite{Fish1995}. Lastly the RL method does not provided an associated estimate of the uncertainties of the reconstruction, which is desirable especially in the Poisson domain to measure the significance of specific features found in the reconstructed image.

@@

    To address the issue with RL decomposing extended emission structures into groups of point source, multiple additions to the RL method were proposed. Most importantly this lead to the introduction of regularised RL methods, such as use of total variation \citep{Dey2006} or sparseness \todo{add reference}. The challenge is due to fundamentally contradicting goals: deconvolution favoring point sources will decompose extend structures, while introducing local correlations will lead to less sharp reconstrcution of point sources.

@@

    \cite{Pumpe2018} and \cite{Selig2015} approached the problem of deconvolution of Poisson noise data implicitly using the framework of information field theory. Deconvolution, Denoising and Decomposition (D3PO) and (D4PO).

    Recently, image deconvolution using physical assumptions and decomposition into point and diffuse flux in 3D/4D has gained attention. proposed priors based on physical assumptions, which have been shown to work well for large-scale emission. However, the smoothness assumption in these methods can lead to artifacts in the galactic plane.

    Interesting approaches to the problem of deconvolution can of course be found in the field of computer vision. For example \cite{Zoran2011} build on very similar assumption of local correlation for natural images. Instead of assuming a fixed local correlation structure, such as smoothness defined by a Gaussian with a correlation size, they proposed to learn the local correlation structure on millions of patches extracted from a set of reference images. This approach can be seen as a form of transfer learning. For this they proposed to split training images from representative data sets into small patches of size 8x8~pixels and fit a 64 dimensional Gaussian Mixture Model (GMM) to the distribution of extracted patches. Each pixel of the patch 
    is thereby treated as an independent dimension in the model. In total
    they chose to use $k=200$ Gaussian components. To use the GMM in reconstruction they introduce the expected patch likelihood (EPLL). They showed that the patch based GMM prior led to much improved image reconstructions for multiple inverse problems such as denoising, deblurring and inpainting. \cite{Bouman2016} later adapted the patch prior successfully for the reconstruction of radio data from the Event Horizon Telescope (EHT).

@@

    Some of the challenges of deblurring with CNN are due to the convolutional nature of the networks. In a single layer, linear approximation, the network would need to learn the inverse filter of the PSF, corresponding to the well known \textit{Wiener filter} solution of the problem. This requires large convolution kernels or deep network architectures, which require in turn large amounts of training data. This challenge was described early by \cite{Li2014} who introduced a baseline architecture for a deep CNN for image deblurring.

    One last important challenge for deep learning based methods is obtaining uncertainty estimates for the reconstructed image from CNNs. In the limit of Poisson noise it is desirable to trace the full a-postiori likelihood, which is not possible for most network architectures, with the exception of for example normalising flows.
    This problem could be adresses with Monte Carlo base methods, such as bootstrapping, further increasing the demand on training data and computation ressources.

@@

    For each individual observation $j$ the predicted counts can be modelled by \textit{forward folding} the unknown flux image $\mathbf{x}$ through the individual instrument response:
    %
    \begin{equation}
        \label{eq:model}
        \boldsymbol{\lambda}_j(\mathbf{x}) = \mathrm{PSF}_j \circledast \left(\mathbf{E}_j \cdot (\mathbf{x} + \mathbf{B}_j) \right)
    \end{equation}

@@

    Given a single observation $j$ of an unknown flux image
    $\mathbf{x}$ and assuming the noise in each pixel $i$ in the recorded counts image
    $\mathbf{D_j}$ follows a Poisson distribution, the likelihood $\mathcal{P}$
    of obtaining the measured image from a model image of the expected
    counts $\boldsymbol{\lambda_j}$ with $N$ pixels is given by:
    %
    \begin{equation}
        \label{eq:poisson}
        P\left( \mathbf{D_j} | \boldsymbol{\lambda_j} \right) = \prod_{i=1}^N \frac{{e^{ - D_{j,i} } \lambda_{i,j} ^ {D_{j,i}}}}{{D_{j,i}!}}
    \end{equation}

    By taking the logarithm and dropping the constant terms one can transform the
    product into a sum over pixels, which is also often called the \textit{Cash}
    \citep{Cash1979} fit statistics:
    %
    \begin{equation}
        \label{eq:cash}
        \mathcal{C}\left( \mathbf{D_j} | \boldsymbol{\lambda_j} \right) = \frac{1}{N}\sum_{i=1}^N \lambda_{j,i} - D_{j, i} \log{\lambda_{j,i}}
    \end{equation}

    To account for the possibility of different datasets having a different number of pixels we divide by $N$ (\todo{Does this need more justification...?}).
    By summing over all observations we finally get the joint log-likelihood of measuring a set of counts images $\mathbf{D}$ given an unknown flux image $\mathbf{x}$:
    %
    \begin{equation}
        \label{eq:joint}
        \mathcal{L}\left( \mathbf{D} | \mathbf{x} \right) = \sum_{j=1}^J \mathcal{C}\left( \mathbf{D_j} | \mathbf{x} \right)
    \end{equation}

@@

    Using Bayes rule we can estimate the \aposteriori likelihood under a given prior:
    %
    \begin{equation}
        \label{eq:bayes}
        P(\mathbf{x}|\textbf{D}) = P(\mathbf{x} ) \frac{P(\textbf{D} |\mathbf{x})}{P(\textbf{D})}
    \end{equation}

    And use a Maximum A Posteriori (MAP) approach to find an estimate for $\mathbf{x}$. Taking the logarithm of Eq.~\ref{eq:bayes}, replacing the definition of the likelihood $P(\mathbf{D}|\mathbf{x})$ with Eq.~\ref{eq:joint} and dropping the normalisation term $P(\mathbf{D})$ which is independent of $\mathbf{x}$, we get the following expression for the log-posterior likelihood $\mathcal{L}$ :
    %
    \begin{equation}
        \label{eq:total}
        \mathcal{L}\left(\mathbf{x} | \mathbf{D} \right) = \sum_{j=1}^J \mathcal{C}\left( \mathbf{D}_j | \mathbf{x} \right) - \beta \cdot \mathcal{P}(\mathbf{x})
    \end{equation}

@@

    \subsection{Patch Based Image Prior}
    In general it is difficult to capture global image statistics in a prior likelihood taking into account all pixel to pixel correlations. The size of the corresponding correlation matrix would grow with the number of pixels squared and thus quickly becomes computationally untraceable. However astronomical images typically only show local correlations between pixels. This is because the spatial scale of astrophysical processes in an image is always limited by the finite interaction time and the fact that the interaction cannot propagate faster than the speed of light (structure in the CMB might be an exception here). This means on smaller scales astronomical images often contain basic structures such as edges, corners, filaments or periodic patterns formed by local astrophysical processes. For this reason we adopt the idea of a patch based image prior approach as introduced by \cite{Zoran2011}. Given a pre-trained Gaussian Mixture Model (GMM) for the distribution of pixel values in patches, we can evaluate a patch based prior for the reconstructed image $\mathbf{x}$ and approximate the \aposteriori likelihood.

    One convenient property of the GMM is the possibility to evaluate its likelihood in closed form. For a given single patch $\mathbf{x}_n$ it is given by:
    %
    \begin{equation}
        \label{eq:gmm}
        P_{GMM}(\mathbf{x}_n) = \sum_{k=1}^K \pi_k \mathcal{N}(\mathbf{x}_n| \boldsymbol{\mu}_k, \boldsymbol{\Sigma}_k^2)
    \end{equation}

@@

    From Equation~\ref{eq:gmm} we can also derive the log-likelihood for a specific choice of the component $k$:
    %
    \begin{equation}
    \begin{split}
    \mathcal{P}_{GMM}(k| \mathbf{x}_n) = -\frac{1}{2} \left[ \right. d \log(2\pi)
    + \log(|\boldsymbol \Sigma_{k}|)\\
    + (\mathbf{x}_n - \boldsymbol{\mu}_{k})^T \Sigma^{-1}_{k}(\mathbf{x}_n - \boldsymbol{\mu}_{k}) \left. \right]
    \end{split}
    \end{equation}

    If we split the image into non-overlapping patches, we can find for each noisy patch $\mathbf{x}_n$ the GMM component $\hat{k}_n$, which maximises the log-likelihood of the GMM:
    %
    \begin{equation}
        \hat{k}_n = \underset{k \in \{1, ..., N\} }{\argmax}{\mathcal{P}_{GMM}(k| \mathbf{x}_n)}
    \end{equation}
    \vspace{0.2em}

    As the number of components in the GMM is limited we can do this simply by evaluating the likelihood for each of the components and choose the one with the maximum value of the log-likelihood. To finally evaluate the log-likelihood of the total image we can sum up the maximum log-likelihood values of the individual patches: 
    %
    \begin{equation}
        \mathcal{P}(\mathbf{x}, \mathbf{k}) = \sum_{n = 1}^N \mathcal{P}_{GMM}(\hat{k}_n | \mathbf{P}_n \mathbf{x})
    \end{equation}

@@

    In the case of non-overlapping patches the

@@

    To account for both systematic effects we extend the model for the predicted counts introduced in Eq.~\ref{eq:model}, by a background normalisation factor $\alpha_j$ and a linear shift in position $\phi_j(\mathbf{x}| \delta_{x,j}, \delta_{y,j})$, which is specific for each observation. This yields the following modified model definition:
    %
    \begin{equation}
        \label{eq:model-npred-calibration}
        \mathbf{\lambda}_j = \mathrm{PSF}_j \circledast \left(\mathbf{E}_j \cdot (\phi_j(\mathbf{x}| \delta_{x,j}, \delta_{y,j}) + \alpha_j \cdot \mathbf{B}_j) \right)
    \end{equation}

@@

    \todo{One could try to estimate a sparse covariance matrix...or approximate it using the GMM covariance...}

@@

    We implemented the \jolideco method as an independent Python package, based on \texttt{PyTorch} as optimization back-end. We tried to define a modular, object oriented code structure, to allow parts of the algorithm to be flexible and interchangeable, such as choice of the patch normalisation, the choice of GMM models, optimization methods and serialisation formats for the reconstruction results as well as corresponding diagnostic information such as the trace of the posterior and prior likelihood values. We hope that this simplifies the future improvement and extension of the package. The package is available at \url{https://github.com/jolideco/jolideco}.

@@

    \subsubsection{JWST Cassiopeia A}
    The James Webb Space Telescope (JWST) is a large, infrared space telescope launched in late 2021. It delivers image data a very high spatial resolution as well as excellent signal to noise ratio. Both of these properties make JWST data an excellent choice for extracting patches for training a GMM. To limit the data reduction effort we relied on a promotional image of the SNR \textit{Cassiopeia A}\footnote{The image was downloaded from \url{https://webbtelescope.org/contents/media/images/2023/121/01GWQBBY77MHGFV3M3N63KDCEJ}}.
    We first converted the image to gray scale using a perceived brightness model \todo{reference skimage...}. Then we split the image into $approx 2$ million patches and learned a GMM with 128 mixture components.

@@

    \begin{figure*}
        \script{example-chandra.py}
        \begin{centering}
            \includegraphics[width=\linewidth]{figures/example-chandra.pdf}
            \caption{
               The \jolideco method applied to \chandra data of the SNR \textit{1E~0102.2–7219}. The left large image shows the summed counts from 25 observations. The right large image  shows the result from the \jolideco deconvolution process. The flux is given as counts relative to the reference observation with ID=1308. The panels \textit{Zoom A}, \textit{Zoom B} and \textit{Zoom C} show a zoomed-in version of three example regions in the SNR, marked with the white rectangles in the large images. The colormap and scale used in the zoom insets is the same as used for the larger images.
            }
            \label{fig:example-chandra}
        \end{centering}
    \end{figure*}

@@

    The \texttt{fermitools} output an energy dependent counts, exposure and PSF cube. To further reduce the data to images we use the \textit{Gammapy} package v1.0 \citep{GammapyZenodov1.0.1, Donath2023}. To compute an image of the expected background emission we rely on the latest Galactic diffuse emission model \texttt{gll\_iem\_v07.fits} and also include nearby sources from the 3FHL catalog \cite{Ajello2017}. We first fit the model components to the data using \textit{Gammapy}, excluding a circular region of radius \qty[mode = text]{1.02}{$\deg$} centered on \textit{Vela Junior}. Finally we reduce this model to an image by summing over the energy dimension. To compute the exposure and PSF image we assume a power law with index $\Gamma=1.77$ \citep{Ajello2017} as a spectral model for \textit{Vela Junior} and integrate over the energy dimension of the cube while weighting with the spectral model.

@@

    \section{Paper Reproducibility}
    \label{sec:reproducibility}
    In order to facilitate reproducibility of our results we provide the source code for the paper and figures at \url{https://github.com/jolideco/jolideco-paper}. The paper can be reproduced using using the tool \texttt{Showyourwork} \citep{Luger2021} and following the instructions in the \texttt{README.md} file in the same repository. All data is available for download from Zenodo \todo{how do we do this exactly...?} or can be reproduced using the custom  \texttt{Snakemake} workflows and repositories provided under the \jolideco GitHub organisation\footnote{\url{https://github.com/jolideco}}. For the Fermi-LAT data reduction we used a custom \texttt{Snakemake} workflow  available at\footnote{\url{https://github.com/adonath/snakemake-workflow-fermi-lat}}. For the Chandra data reduction we used a custom \texttt{Snakemake} workflow available at\footnote{\url{https://github.com/adonath/snakemake-workflow-chandra}}.

@@

    To increase the size of the correlation window of the patch prior, \cite{Papyan2015} proposed to evaluate and combine the prior on patches extracted at multiple scales of the image. This is a promising approach to further improve the reconstruction quality for extended sources with low signal to noise ratio.

    To further improve the general expressiveness of the prior \cite{Altekrueger2022} used normalising flows (NF) to model the distribution of patches in the context of computer tomography image reconstruction. Just like a GMM the NF allows to evaluate its likelihood function and therfore construct an explicit, approximate \aposteriori probability for optimization. \todo{However it is also know that NFs struggle with multi-modality...not sure how this affects the propsoed approach...}

    Another promising approach to handle prior information for images has been introduced by \citep{Ulyanov2017} with the so called \enquote{Deep Image Prior}. They showed that a prior structure for an image can be built into a deep neural network, where the architecture of the network encodes the prior information on the correlation structure between neighbouring pixels. The authors showed, that the deep image prior successfully imposes self similarity on the structures in the image. As structures in astronomical images often show sell-similarity this is and interesting approach, that could even be combined with the patch prior.
